An Analysis of Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters

### 2010-10-21 doc/isbi/thesis.tex
more on isbi

\usepackage[printonlyused,withpage]{acronym}
\usepackage{colortbl}
\usepackage{times}
\usepackage[siunitx]{circuitikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cancel}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}

@@

Traditional methods of analyzing functional Magnetic Resonance 
Images use a linear combination of
just a few static regressors. This work demonstrates an alternative
approach using a physiologically inspired nonlinear model. By using a 
particle filter to optimize the model parameters, the computation time
is kept below a minute per voxel without requiring a linearization 
of the noise in the state
variables. The activation results show regions similar to those found in 
Statistical Parametric Mapping; however, there are some notable regions not 
detected by that technique. Though the parameters selected by the particle filter based approach
are more than sufficient to predict the Blood-Oxygen-Level-Dependent signal
response,
more model constraints are needed to uniquely identify a single set
of parameters. This ill-posed nature explains the large discrepancies
found in other research that attempted to characterize the model parameters.
For this reason the final distribution of parameters is more medically relevant
than a single estimate. Because the output of the particle filter is 
a full posterior probability, the reliance on the mean to estimate 
parameters is unnecessary. This work presents
not just a viable alternative to the traditional method of detecting
activation, but an extensible technique of estimating the joint probability
distribution function of the Blood-Oxygen-Level-Dependent Signal parameters.

@@

The Neuroscientists at Wake Forest have also been extremely helpful. In 
particular I would like to thank Paul Laurienti for help gathering fMRI
data and using SPM.

To my parents, for all the psychological (and financial) help along the way,
thank you so much. You were always there for me, I couldn't
ask for better parents. Love you guys.

@@

\huge
\bfseries
List of Acronyms
\mdseries
\normalsize
\begin{acronym}[CMRO2]
\acro{BOLD}{Blood-Oxygen-Level-Dependent}
\acro{CBF}{Cerebral Blood Flow}
\acro{CBV}{Cerebral Blood Volume}
\acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
\acro{DC}{Direct Current}
\acro{dHb}{Deoxygenated Hemoglobin}
\acro{EM}{Expectation-Maximization}
\acro{EPI}{Echo Planar Imaging}
\acro{fMRI}{Functional Magnetic Resonance Imaging}
\acro{FMRIB}{Oxford Centre for Functional MRI of the Brain}
\acro{FSL}{FMRIB Software Library}
\acro{FWHM}{Full-Width Half-Maximum}
\acro{GA}{Genetic Algorithms}
\acro{GLM}{General Linear Model}
\acro{HRF}{Hemodynamic Response Function}
\acro{Hb}{Hemoglobin}
\acro{MAD}{Median Absolute Deviation}
\acro{MI}{Mutual Information}
\acro{MR}{Magnetic Resonance}
\acro{MRI}{Magnetic Resonance Imaging}
\acro{MSE}{Mean Squared Error}
\acro{O2Hb}{Oxygenated Hemoglobin}
\acro{ODE}{Ordinary Differential Equation}
\acro{PDF}{Probability Density Function}
\acro{POSSUM}{Physics-Oriented Simulated Scanner for Understanding MRI}
\acro{RF}{Radio Frequency}
\acro{RMSE}{Root Mean Squared Error}
\acro{RMSR}{Root Mean Squared Residual}
\acro{SA}{Simulated Annealing}
\acro{SNR}{Signal-to-Noise Ratio}
\acro{SPM}{Statistical Parametric Mapping}
\acro{T1}{Longitudinal}
\acro{T2}{Spin-Spin}
\acro{TR}{Repetition Time}
\acro{UKF}{Unscented Kalman Filter}
\end{acronym}
\pagebreak

\pagenumbering{arabic}
\pagestyle{myheadings}
\acresetall
\include{intro}
\include{background}
\include{particlefilters}
\include{methods}
\include{results}
\include{resultsB}
\include{discussion}
%\include{conclusion}

>>>>

\usepackage{colortbl}
\usepackage{times}
\usepackage[siunitx]{circuitikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cancel}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}

@@

Traditional methods of analyzing FMRI images use a linear combination of
just a few static regressors. This work demonstrates an alternative
approach using a physiologically inspired nonlinear model. By using a 
particle filter to optimize the model parameters, the computation time
is kept below a minute per voxel without requiring a linearization 
of the noise in the state
variables. The activation results show regions similar to those found in 
SPM; however, there are some notable regions not detected by 
SPM. Though the parameters selected by the particle filter based approach
are more than sufficient to predict the BOLD response,
more model constraints are needed to uniquely identify a single set
of parameters. This ill-posed nature explains the large discrepancies
found in other research that attempted to characterize the model parameters.
For this reason the final distribution of parameters is more medically relevant
than a single estimate. Because the output of the particle filter is 
a full posterior probability, the reliance on the mean to estimate 
parameters is unnecessary. This work presents
not just a viable alternative to the traditional method of detecting
activation, but an extensible technique of estimating the joint probability
distribution function of the BOLD parameters.

@@

To my parents, for all the psychological (and financial) help along the way,
thank you so much. When things were the darkest you were there for me, I couldn't
ask for better parents. Love you guys.

@@

\pagenumbering{arabic}
\pagestyle{myheadings}

\include{intro}
\include{background}
\include{particlefilters}
\include{methods}
\include{results}
\include{resultsB}
\include{discussion}
\include{conclusion}
